---
title: "Chapter 1"
author: "B Davenport"
date: "22/04/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
```

## 1.1 

```{r A first map} 
system.file("gpkg/nc.gpkg", package = "sf") %>% 
    read_sf() -> nc
nc.32119 <- st_transform(nc, 32119)
nc.32119 %>% 
    select(BIR74) %>% 
    plot(graticule = TRUE, axes = TRUE)
```

```{r data for 1.1}
nc %>% select(AREA, BIR74, SID74) %>% print(n = 3)
```
- the (selected) dataset has 100 features (records) and 3 fields (attributes)
- the geometry type is MULTIPOLYGON (chapter 3)
- it has dimension XY, indicating that each point will consist of 2 coordinate values
- the range of x and y values of the geometry
- the coordinate reference system (CRS) is geodetic, with coordinates in degrees longitude and latitude associated to the NAD27 datum (chapter 2)
- the three selected attribute variables are followed by a variable geom of type MULTIPOLYGON with unit degrees that contains the polygon information

```{r facet plots}
year_labels = c("SID74" = "1974 - 1978", "SID79" = "1979 - 1984")
nc.32119 %>% select(SID74,SID79) %>% gather(VAR, SID, -geom) -> nc2
ggplot() + geom_sf(data = nc2, aes(fill = SID)) +
    facet_wrap(~VAR, ncol = 1, labeller = labeller(VAR = year_labels)) +
    scale_y_continuous(breaks = 34.36) +
    scale_fill_gradientn(colors = sf.colors(20)) +
    theme(panel.grid.major = element_line(color = "white"))
```

```{r an interactive map}
suppressPackageStartupMessages(library(mapview))
nc.32119 %>% mapview(zcol = "BIR74", legend = TRUE, col.regions = sf.colors)
```

